---
title: "PML"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Practical Maching Learning Project
## Biqing Sheng

## Executive Summary
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Three types of classification techniques were used for prediction: decision tree, random forest, and generalized boosting model. The accuracy of each model was evaluated. The model with the best accuracy was used to predict the test data set.

## Data Preparation
```{r preparation}
# Loading required packages
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(e1071)
# Read in data
pml_training <- read.csv("pml-training.csv")
pml_testing <- read.csv("pml-testing.csv")
# The first 7 variables were eliminated since there effects are minimal
training1 <- pml_training[,-c(1:7)]
testing1 <- pml_testing[,-c(1:7)]
# Remove the column with missing values
training2 <- training1[, colSums(is.na(training1)) == 0]
testing2 <- testing1[, colSums(is.na(training1)) == 0]
# Remove the variables with near zero variance
NZV <- nearZeroVar(training2)
training2 <- training2[,-NZV]
dim(training2)
```

The original traing data set was seperated randomly. 60% was used for training, 40% was used for validation.
```{r dataseperation}
set.seed(121)
inTrain <- createDataPartition(training2$classe, p = 0.6, list = FALSE)
traindata <- training2[inTrain,]
validdata <- training2[-inTrain,]
dim(traindata)
```

## Decision Tree Model
```{r DTmodel}
set.seed(121)
model_DT <- rpart(classe~., data = traindata, method = "class")
fancyRpartPlot(model_DT)
prediction1 <- predict(model_DT, validdata, type = "class")
cm1 <- confusionMatrix(prediction1,validdata$classe)
```
The accuracy of decision tree model is `r cm1$overall[1]`.

## Random Forest Model
```{r RFmodel}
set.seed(121)
trctrl <- trainControl(method = "cv", number = 5)
model_RF <- train(classe~., data = traindata, method = "rf", trControl = trctrl, verbose=FALSE)
plot(model_RF$finalModel)
prediction2 <- predict(model_RF,validdata)
cm2 <- confusionMatrix(prediction2,validdata$classe)
```
The accuracy of random forest model with 5-fold cross-validation is `r cm2$overall[1]`.

## Generalized Boosting Model
```{r GBMmodel}
set.seed(121)
trctrl <- trainControl(method = "cv", number = 5)
model_GBM <- train(classe~., data = traindata, method = "gbm", trControl = trctrl, verbose = FALSE)
plot(model_GBM)
prediction3 <- predict(model_GBM,validdata)
cm3 <- confusionMatrix(prediction3,validdata$classe)
```
The accuracy of generalized boosting model with 5-fold cross-validation is `r cm3$overall[1]`.

## Final Prediction for Test Data
```{r finalprediction}
predict(model_RF,testing2)
```

